Assessing the Replicability of Psychological Science Through Pedagogy

### 2016-09-08 paper/paper.Rmd
bib updates

Past work has attempted to estimate the proportion of true effects in a particular literature [@ioannidis2005;@osc2015], but this goal is challenging -- perhaps even impossible [@gilbert2016;@anderson2016]. Assessing the truth of even an individual effect relies on delineating what constitutes an exact replication [@cites] and establishing a standard for replication success [@simonsohn2015]. Estimating the proportion of true effects in a literature further requires a systematic sampling strategy to ensure generality. Furthermore, no single replication attempt is definitive, so multiple iterations are typically necessary to increase the evidential value of replication attempts [@lewis2016;@klein2014].

@@

```{r}
ppt <- project_plan %>%
        select(cite, expt_num, open_data,
               open_materials, original_on_turk, original_n,collected_n)

names(ppt) = c("Citation", "Expt", "Data", "Materials", "Turk?", 
               "N (orig)","N (rep)")

# print(xtable(ppt, caption = "Summary of attempted replications", 
#        label = "tab:summary"), include.rownames=FALSE)
```

@@

Atir, Rosenzweig, \& Dunning & 1b & No & Yes & Yes & 202 & Other &  50 & 6.50 & Full \\ 
Ko, Sadler, \& Galinksy & 2 & Yes & Some & No &  40 & Original &  40 & 4.50 & Full \\ 
Lewis \& Oyserman & 4 & No & Yes & Yes & 122 & 80\% power & 128 & 6.25 & Partial \\ 
Liverence \& Scholl & 1 & No & Some & No &  18 & Original &  19 & 4.75 & Partial \\ 
Proudfoot, Kay, \& Koval & 1 & No & No & Yes &  80 & 80\% power &  84 & 6.25 & Full \\ 
Scopelliti, Loewenstein, \& Vosgerau & 3 & No & Yes & Yes & 550 & 80\% power & 124 & 4.75 & None \\ 
Sofer et al. & 1 & No & Yes & No &  48 & Other &  95 & 4.50 & Partial \\ 
Storm \& Stone & 3 & No & No & No &  48 & Original &  61 & 4.00 & Partial \\ 
Wang et al. & 2 & No & No & No & 219 & 80\% power & 397 & 4.00 & None \\ 
Xu \& Franconeri & 1a & No & No & No &  12 & Original &  27 & 4.00 & Partial \\ 
Zaval, Markowitz, \& Weber & 1 & Yes & Yes & Yes & 312 & Original & 321 & 5.25 & Partial \\ 
   \hline
\end{tabular}
\caption{\label{tab:summary} Summary of original research papers and results of experiments that were targets for replication in our projects.} 
\label{tab:summary}
\end{table}

@@

All study participants were recruited on Amazon Mechanical Turk; individual samples are given in Table \ref{tab:summary}.^[Note that of the `r nrow(ppt)` studies included in the final sample, `r sum(project_plan$original_on_turk=="Yes")` (`r 100 * round(mean(project_plan$original_on_turk=="Yes"), digits=2)`%) were originally conducted on Mechanical Turk.] All experiments were approved by the Stanford University institutional review board under protocol \#23274, "Reproducibility of psychological science and instruction."

@@

Of the `r nrow(ppt)` studies included in the final sample, `r 100 * round(mean(project_plan$open_materials == "Yes" | project_plan$open_materials == "Some"), digits=2)`% of studies had either some or all materials openly available. For `r sum(project_plan$author_contact != "")` studies, however, students contacted the authors to request either materials or clarification of methods (using a template email that was customized by the students and reviewed by the instructors). Responses were received in all but one case, all within a matter of days.

@@

Since our sample only includes 11 studies, we describe general trends here rather than attempting to conduct statistical tests (which would be dramatically under-powered after correcting for multiple comparisons). Overall, projects had a somewhat larger probability of being judged successful if they were judged to be closer to the original (`r 100 * round(mean(d$replicated_TA[d$high_fidelity]), digits=2)`\% 
vs. 
`r 100 * round(mean(d$replicated_TA[d$high_fidelity == FALSE]), digits=2)`\%). This result is consistent with a reanalysis of @osc2015 that found identified variability across contexts as a potential factor mediating replication success [@vanbavel2016; cf. @inbar2016].
There were also small numerical effects of being run on Mechanical Turk (`r 100 *round(mean(d$replicated_TA[d$original_on_turk == "Yes"]), digits=2)`\% vs. `r 100 *round(mean(d$replicated_TA[d$original_on_turk == "No"]), digits=2)`\%) and having open materials (`r 100 *round(mean(d$replicated_TA[d$open_materials!="No"]), digits=2)`\% 
vs. `r 100 *round(mean(d$replicated_TA[d$open_materials=="No"]), digits=2)`\%), though we do not believe these differences are interpretable given our sample size.

>>>>

Past work has attempted to estimate the proportion of true effects in a particular literature [@ioannidis2005;@osc2015], but this goal is challenging -- perhaps even impossible [@gilbert2016;@anderson2016]. Assessing the truth of even an individual effect relies on delineating what constitutes an exact replication [e.g., @stroebe2014] and establishing a standard for replication success [e.g., @simonsohn2015]. Estimating the proportion of true effects in a literature further requires a systematic sampling strategy to ensure generality. Furthermore, no single replication attempt is definitive, so multiple iterations are typically necessary to increase the evidential value of replication attempts [@lewis2016;@klein2014].

@@

@@

Atir, Rosenzweig, \& Dunning & 1b & No & Yes & Yes & 202 & Other &  50 & 6.67 & Full \\ 
Ko, Sadler, \& Galinksy & 2 & Yes & Some & No &  40 & Original &  40 & 4.67 & Full \\ 
Lewis \& Oyserman & 4 & No & Yes & Yes & 122 & 80\% power & 128 & 6.67 & Partial \\ 
Liverence \& Scholl & 1 & No & Some & No &  18 & Original &  19 & 5.33 & Partial \\ 
Proudfoot, Kay, \& Koval & 1 & No & No & Yes &  80 & 80\% power &  84 & 6.67 & Full \\ 
Scopelliti, Loewenstein, \& Vosgerau & 3 & No & Yes & Yes & 550 & 80\% power & 124 & 5.67 & None \\ 
Sofer et al. & 1 & No & Yes & No &  48 & Other &  95 & 4.33 & Partial \\ 
Storm \& Stone & 3 & No & No & No &  48 & Original &  61 & 4.00 & Partial \\ 
Wang et al. & 2 & No & No & No & 219 & 80\% power & 397 & 4.33 & None \\ 
Xu \& Franconeri & 1a & No & No & No &  12 & Original &  27 & 5.00 & Partial \\ 
Zaval, Markowitz, \& Weber & 1 & Yes & Yes & Yes & 312 & Original & 321 & 5.00 & Partial \\ 
   \hline
\end{tabular}
\caption{\label{tab:summary} Summary of original research papers and results of experiments that were targets for replication in our projects.} 
\label{tab:summary}
\end{table}

@@

All study participants were recruited on Amazon Mechanical Turk; individual samples are given in Table \ref{tab:summary}.^[Note that of the `r nrow(d)` studies included in the final sample, `r sum(project_plan$original_on_turk=="Yes")` (`r 100 * round(mean(project_plan$original_on_turk=="Yes"), digits=2)`%) were originally conducted on Mechanical Turk.] All experiments were approved by the Stanford University institutional review board under protocol \#23274, "Reproducibility of psychological science and instruction."

@@

Of the `r nrow(d)` studies included in the final sample, `r 100 * round(mean(project_plan$open_materials == "Yes" | project_plan$open_materials == "Some"), digits=2)`% of studies had either some or all materials openly available. For `r sum(project_plan$author_contact != "")` studies, however, students contacted the authors to request either materials or clarification of methods (using a template email that was customized by the students and reviewed by the instructors). Responses were received in all but one case, all within a matter of days.

@@

Since our sample only includes 11 studies, we describe general trends here rather than attempting to conduct statistical tests (which would be dramatically under-powered after correcting for multiple comparisons). Overall, projects had a slightly larger probability of being judged successful if they were judged to be closer to the original (`r 100 * round(mean(d$replicated_TA[d$high_fidelity == TRUE]), digits=2)`\% 
vs. 
`r 100 * round(mean(d$replicated_TA[d$high_fidelity == FALSE]), digits=2)`\%). 
There were also small numerical effects of being run on Mechanical Turk (`r 100 *round(mean(d$replicated_TA[d$original_on_turk == "Yes"]), digits=2)`\% vs. `r 100 *round(mean(d$replicated_TA[d$original_on_turk == "No"]), digits=2)`\%) and having open materials (`r 100 *round(mean(d$replicated_TA[d$open_materials!="No"]), digits=2)`\% 
vs. `r 100 *round(mean(d$replicated_TA[d$open_materials=="No"]), digits=2)`\%), though we do not believe these differences are interpretable given our sample size.
